### 2014-11-10 Arvind.tex
Pass through Decomposing

\begin{figure}
\centering
\includegraphics[scale=.43]{zoom}
\caption{Components of a multi-processors system}
\label{zoom}
\end{figure}

Any conventional multi-processor system can be logically divided into 3 components as shown in Figure \ref{zoom}. 
The top-level system design is shown at top right, while the details of its components, the memory system and the processor ($P_i$), are shown in the magnified boxes. The processor component $P_i$ can be implemented in a variety of ways, from one which executes instructions one-by-one in program order, to a complex one which speculatively executes many instructions concurrently to exploit parallelism. The memory component is normally implemented using a hierarchy of caches, in order to increase the performance of the overall system, since the latency of accessing memory directly is large compared to that of accessing a much smaller cache. The component between each processor and the global memory sub-system contains local buffers, $B_i$, specific to
the processor $P_i$.

@@

Popular instruction set architectures, such as Intel x86, ARM, PowerPC, do not guarantee sequential consistency.  That is, their behavior makes it unsound to imagine that
the different processors of a shared-memory system are running
according to simple interleaving of instructions, where each instruction executes atomically.  However, we want to emphasize that, in every weak-memory system we are aware of, \emph{the main memory still exposes atomic loads and stores}!  The weaker semantics arises only because of (1)
reordering of memory instructions by the core $P_i$ and/or (2) the
properties of the local buffers $B_i$ connected to each processor
$P_i$.  For instance, if $B_i$ were a store buffer, into which the
committed stores from an in-order processor $P_i$ are enqueued, and if
any ``subsequent'' load from the same processor $P_i$ is allowed to read and commit the latest value from the store buffer, then such a multi-processor system is said to implement the TSO~\cite{x86tsocacm10} memory
model.

@@

So, we focus on this opportunity to simplify proof decomposition.  We will
prove that our main memory component satisfies an intuitive \emph{store
atomicity} property.  Despite the potential red herring of the word
``atomicity,'' this specification is appropriate even for implementation of
weaker memory models than SC.  Store atomicity can be understood via the
operational semantics of Figure \ref{$M_m$}, describing an LTS that receives
load and store requests ($\ld$ and $\st$) from processors and sends back load
responses ($\ld\resp$). The transfer happens via input buffers $\inp(p)$ from
processor $p$ and output buffers $\outp(p)$ to processor $p$.

The memory component is composed of a \emph{hierarchy of caches}, where each
cache node (labeled like ``L1,'' ``L2,'' etc.) communicates only with its
neighbors in the graph.  Each processor has a dedicated L1 cache, from which we
do our best to satisfy all memory requests, to avoid the latency of a
round-trip with main memory.  With an L1 cache miss, we may still find a hit in
a parent cache below the level of main memory, realizing a smaller but still
significant speedup.  Therefore it is the responsibility of the hierarchy of
caches, which forms the memory sub-component, to implement the store atomicity
property. In fact, as we will prove in Section \ref{sec:cc}, the purpose of the
cache-coherence protocol is to establish this invariant for the memory
sub-component.  Concretely, we have verified a \emph{directory-based} protocol
for coordinating an arbitrary tree of caches, where each node stores a
conservative approximation of its children's states.

As an instance of the above decomposition, we will prove that a multi-processor
system with no local buffering in between the processor and the memory
components indeed implements sequential consistency. We implement a highly
speculative processor that executes instructions and issues loads out of order,
but commits instructions (once some ``verification'' is done) in order. This idea
of verification loads for a speculative processor is not new \cite{cain}.

>>>>

\begin{figure}
\centering
\includegraphics[scale=.43]{zoom}
\caption{Components of a multiprocessor system}
\label{zoom}
\end{figure}

Any conventional multiprocessor system can be divided logically into 3 components as shown in Figure \ref{zoom}. 
The top-level system design is shown at top right, while the details of its components, the memory system and the processor ($P_i$), are shown in the magnified boxes. The processor component $P_i$ can be implemented in a variety of ways, from one executing instructions one-by-one in program order, to a complex one speculatively executing many instructions concurrently to exploit parallelism. The memory component is normally implemented using a hierarchy of caches, in order to increase the performance of the overall system, since the latency of accessing memory directly is large compared to that of accessing a much smaller cache. The component between each processor and the global memory subsystem contains local load/store buffers, $B_i$, each specific to a processor $P_i$.

@@

Popular instruction set architectures, such as Intel x86, ARM, and PowerPC, do not guarantee sequential consistency.  That is, their behavior makes it unsound to imagine that
the different processors of a shared-memory system are running
according to simple interleaving of instructions, where each instruction executes atomically.  However, we want to emphasize that, in every weak-memory system we are aware of, \emph{the main memory still exposes atomic loads and stores}!  The weaker semantics arises only because of (1)
reordering of memory instructions by the core $P_i$ and/or (2) the
properties of the local buffers $B_i$ connected to each processor
$P_i$.  For instance, if $B_i$ were a store buffer, into which the
committed stores from an in-order processor $P_i$ are enqueued, and if
any ``subsequent'' load from the same processor $P_i$ is allowed to read and commit the latest value from the store buffer, then such a multiprocessor system is said to implement the TSO~\cite{x86tsocacm10} memory
model.

@@

So, we focus on this opportunity to simplify proof decomposition.  We will
prove that our main memory component satisfies an intuitive \emph{store
atomicity} property.  Despite the potential red herring of the word
``atomicity,'' this specification is appropriate even for implementation of
weaker memory models than SC.  Store atomicity can be understood via the
operational semantics of Figure \ref{$M_m$}, describing an LTS that receives
load and store requests ($\ld$ and $\st$) from processors and sends back load
responses ($\ld\resp$). The transfer happens via input buffers $\inp(p)$ from
processor $p$ and output buffers $\outp(p)$ to processor $p$.  Note that this
model allows the memory system to answer pending memory requests in any order,
even potentially reordering requests from a single processor,
so long as, whenever it does process a request, that action appears to take
place \emph{atomically}.

The memory component is composed of a \emph{hierarchy of caches}, where each
cache node (labeled like ``L1,'' ``L2,'' etc.) communicates only with its
neighbors in the graph.  Each processor has a dedicated L1 cache, from which we
do our best to satisfy all memory requests, to avoid the latency of a
round-trip with main memory.  With an L1 cache miss, we may still find a hit in
a parent cache below the level of main memory, realizing a smaller but still
significant speedup.  Therefore it is the responsibility of the hierarchy of
caches, which forms the memory subcomponent, to implement the store atomicity
property. In fact, as we will prove in Section \ref{sec:cc}, the purpose of the
cache-coherence protocol is to establish this invariant for the memory
subcomponent.  Concretely, we have verified a \emph{directory-based} protocol
for coordinating an arbitrary tree of caches, where each node stores a
conservative approximation of its children's states.

As an instance of the above decomposition, we will prove that a multiprocessor
system with no local buffering in between the processor and the memory
components indeed implements sequential consistency. We implement a highly
speculative processor that executes instructions and issues loads out of order,
but commits instructions (once some ``verification'' is done) in order. This idea
of verification loads for a speculative processor is not new \cite{cain}.
